# NSF-funded projects on Databrary

This page summarizes data about NSF-funded projects that have used Databrary
to store and share data.

## Background  {-}

Rick Gilmore and Andrea Seisler developed code to generate a series of reports 
about Databrary's users, assets, and data.

In summer 2023, that code was migrated to the Databrary organization on GitHub.
The code base for these reports can be found at <https://github.com/databrary/analytics/>.
The latest analytics report can be found at <https://databrary.github.io/analytics>.

**Note**: Databrary's API does not permit access to datasets or volumes that have yet to be shared by the volume owner.
This means that the data we report here are an underestimate of the total number of projects and datasets that have actually received NSF funding.

## Volume funders {-}

Databrary permits users to list funders associated with a dataset.
The funding information across the site is updated regularly.
The latest report can be found at <https://databrary.github.io/analytics/funders.html>.

As of 2023-07-24, NSF was listed as a funder by the largest number of volumes on Databrary.

Here, we regenerate the data.

### Setup {-}

Most of the work is done via functions listed in `R/functions.R`.

```{r echo=TRUE}
source("R/functions.R")
```

### Databrary volumes that mention NSF {-}

To query Databrary for the volumes that mention an NSF award, we run the following.

As of 2023-07-24, volume 1627 was the largest volume index.
So, we can query Databrary volume-by-volume, to collect data about volumes that list NSF as a funder, as follows:

```{r, eval=FALSE, message=FALSE, warning=FALSE, echo=TRUE}
db_w_nsf <-
  purrr::map(1:1627,
             get_vol_awards,
             funder_str = "NSF",
             .progress = "Databrary vols w/ NSF:") |> purrr::list_rbind() |>
  add_clean_nsf_award_id()
```

The above uses the `get_vol_awards()` function and the `funder_str` parameter to select award information that lists NSF. Then, the NSF award IDs are cleaned to match the NSF format with `add_clean_nsf_award_id()`.

::: {.rmdnote}

The funding award IDs must be cleaned because the current Databrary system does not validate their formats. 
The funding award information is a free text field.
It would be helpful if a future version of Databrary could validate these entries to reduce *post hoc* data cleaning.

:::

This takes several moments, so for this report, we import a previously saved file with the same data.

```{r import-saved-db-nsf, eval=TRUE, echo=TRUE}
db_nsf <- readr::read_csv("csv/databrary-vols-w-nsf-funding-all.csv", show_col_types = FALSE)
```

Then, we can start to visualize the data. Here is a table with all of the shared Databrary volumes mentioning NSF as a funder:

```{r show-table-databrary-vols-listing-nsf, echo=TRUE}
db_nsf <- db_nsf |>
  dplyr::filter(!is.na(award_id)) |>
  dplyr::mutate(url = paste0("https://databrary.org/volume/", vol_id))
  
db_nsf |>
  dplyr::select(vol_id, award_id, url) |>
  knitr::kable(format = 'html') |>
  kableExtra::kable_material() |>
  kableExtra::scroll_box(width = "700px", height = "400px")
```

::: {.rmdnote}

Note that we have added a URL that links directly to the shared volume on Databrary.

:::

There are $n=$ `r length(unique(db_nsf$vol_id))` unique shared Databrary volumes that list NSF as a funder, and $n=$ `r length(unique(db_nsf$award_id))` unique NSF awards.
Some NSF awards have resulted in multiple shared Databrary volumes.

## NSF funding data on these projects {-}

We can now match these projects with funding information available about them from NSF directly.

### Illustration using the current grant {-}

Before we do that, we will illustrate how the `get_nsf_data_for_award_id()` function can be used to gather information from the NSF funding database about the current project.

```{r echo=TRUE}
nsf_2032713_df <- get_nsf_data_for_award_id(2032713)

nsf_2032713_df
```

### Funding information from shared Databrary volumes {-}

To re-generate a data frame with the Databrary volumes for which we are able to extract more details about the NSF funding from the NSF itself, we run

```{r, eval=FALSE, echo=TRUE}
summarize_mult_vol_nsf_awards(1:1627, vb = TRUE)
```

This takes several minutes, so for this report, we will import a previously saved file with these data.

```{r import-saved, echo=TRUE}
nsf <- readr::read_csv("csv/databrary-nsf-funding-details-all.csv", show_col_types = FALSE)
```

Now, we can begin to visualize this, as well:

```{r show-table-nsf-info, echo=TRUE}
nsf <- nsf |>
   dplyr::filter(!is.na(piLastName), !is.na(piFirstName), !is.na(award_id))
 
nsf |>
  dplyr::select(piLastName, piFirstName, award_id, date, title, vol_id, fundsObligatedAmt) |>
  dplyr::mutate(piLastName = stringr::str_to_title(piLastName),
                piFirstName = stringr::str_to_title(piFirstName)) |>
  dplyr::arrange(piLastName, award_id, vol_id) |>
  knitr::kable(format = 'html') |>
  kableExtra::kable_material() |>
  kableExtra::scroll_box(width = "700px", height = "300px")
```

### Visualizations {-}

```{r nsf-award-histogram-all, fig.cap="Histogram of NSF awards for all Databrary volumes.", echo=TRUE}
nsf |>
  dplyr::mutate(award_amt = as.numeric(fundsObligatedAmt)) |>
  ggplot2::ggplot() +
  ggplot2::aes(award_amt) +
  ggplot2::geom_histogram(bins=10)
```

There are $n=$ `r dim(db_nsf)[1]` Databrary volumes that list NSF as a funder.

These represent $n=$ `r db_nsf$award_id |> na.omit() |> unique() |> length()` unique NSF awards.

Of these, we were able to extract data from the NSF API for $n=$ `r nsf$award_id |> na.omit() |> unique() |> length()` NSF awards.

The following volumes do not have matching NSF awards.

```{r}
db_not_found_in_nsf <- !(db_nsf$award_id %in% nsf$award_id)

db_nsf |>
  dplyr::filter(db_not_found_in_nsf) |>
  dplyr::mutate(url = paste0("https://databrary.org/volume/", vol_id)) |>
  dplyr::select(vol_id, award, url) |>
  knitr::kable(format = 'html') |>
  kableExtra::kable_material() |>
  kableExtra::scroll_box(width = "700px", height = "200px")
```

The NSF award IDs must be in error for these volumes.

We can search the NSF database for the term 'databrary' to see if we can find the correct
NSF award ID.

```{r, echo=TRUE}
nsf_w_databrary <- get_nsf_data_for_keyword("databrary")
```

Curiously, the following projects mention "Databrary":

```{r, echo=TRUE}
nsf_w_databrary |>
  dplyr::select(piFirstName, piLastName, id, title) |>
  dplyr::arrange(piLastName, piFirstName) |>
  knitr::kable(format = 'html') |>
  kableExtra::kable_material() |>
  kableExtra::scroll_box(width = "700px", height = "400px")
```

We can filter these based on the last names of the Databrary volume owners.

```{r, echo=TRUE}
nsf_w_databrary |>
  dplyr::select(piFirstName, piLastName, id) |>
  dplyr::filter(piLastName %in% c("Smith", "Bonawitz")) |>
  dplyr::arrange(id) |>
  knitr::kable(format = 'html') |>
  kableExtra::kable_material() |>
  kableExtra::scroll_box(width = "700px", height = "200px")
```

From this result, it looks like the correct NSF award IDS are `1523982` and possibly `1623486`.
Databrary staff will contact the PIs.

::: {.rmdnote}

Note that this workflow helps evaluate the quality of the data and metadata stored on Databrary.

We are considering incorporating these functions into the `databraryr` package in a future release.

:::